The Galactic Midplane is Not a Plane: Implications for {\em Gaia}

### 2019-04-16 paper/ms.tex
Add AR to table
Update Delta J_i references in paper

\begin{deluxetable*}{ccccccc}
\tablecaption{Description and names of the three orbits considered in this
work.\label{tab:orbits}}
\tablehead{\colhead{name} & \colhead{initial position} & \colhead{initial
velocity} & \colhead{$J_r$} &
\colhead{$L_z$} & \colhead{$J_z$} & \colhead{$z_{\text{max}}$}\\ \colhead{ } &
\colhead{$\mathrm{kpc}$} & \colhead{$\mathrm{km}/\mathrm{s}$} &
\colhead{$\mathrm{kpc\,km\,s^{-1}}$} &
\colhead{$\mathrm{kpc\,km\,s^{-1}}$} & \colhead{$\mathrm{kpc\,km\,s^{-1}}$} &
\colhead{$\mathrm{kpc}$}}
\startdata 
thin-disk & $(8, 0, 0)$ & $(0, -190, 10)$ & 40.3 & -1520 & 0.69 & 0.12 \\
thick-disk & $(8, 0, 0)$ & $(0, -190, 50)$ & 32.5 & -1520 & 23.0 & 0.85 \\ 
halo & $(8, 0, 0)$ & $(0, -190, 190)$ & 32.8 & -1520 & 529.1 & 6.16
\enddata
\end{deluxetable*}

@@

\begin{figure}
\plotone{fig/schmactions_Jz_hist.pdf}
\caption{A histogram of the observed values in $J_z$ for the thick-disk orbit
assuming a $z$ offset of $100\,\pc$. One can see that if the observed $z$
values have a bias (from e.g. an incorrectly computed midplane), then the
induced error distribution in $J_z$ is decidedly non-Gaussian. Therefore, any
sort of error propagation must take this into account.}
\label{fig:Jz_hist}
\end{figure}

@@

Empirically describing the error distribution shown in
Figure~\ref{fig:Jz_hist} is difficult. Of course, Gaussian summary statistics
are not adequate. We therefore elect to measure this error by computing the
95\uth{} percentile minus the 5\uth{} percentile of the distribution of
observed action values. We refer to this quantity as $\Delta J_i$ for each
action value. Because of the bimodality of the error distribution, this
quantity roughly measures the distance between the peaks of the two modes.
Furthermore, this bimodality also implies that $\Delta J_i$ is not very
sensitive to the exact percentiles used. This summary statistic does not
reflect the bias induced when the midplane error is $\gtrsim z_{\text{max}}$.

We now repeat the same procedure as in Figure~\ref{fig:one_orbit_wrong_ref}
but for systematic offsets between $0$ and $500\,\pc$ in the $z$ and $x$
components. In Figure~\ref{fig:many_orbit_wrong_ref}, we report the 95\uth{}
minus 5\uth{} percentile divided by the true action value for the three
different fiducial orbits in Table \ref{tab:orbits}. The thick-disk
(\thickcolor) orbit is the one from Figure~\ref{fig:one_orbit_wrong_ref}, but
we also now consider the effect on the action determined for the thin-disk
(\thincolor) and halo (\halocolor) orbits.

The top row of Figure~\ref{fig:many_orbit_wrong_ref} shows the spread induced
in each action for an offset in the $z$-component (i.e. the Galactic
midplane). In the {\em bottom row} we consider offsets in the $x$ component
(i.e. the Solar radius). The {\em left}, {\em middle}, and {\em right} columns
show the fractional spread in the values computed for $J_z$, $L_z$, and $J_r$,
respectively. We compute the fractional spread $\Delta J_i/J_i$ as the 95\uth{}
minus 5\uth{} percentile of the action computed using the offset coordinate
system over the course of the first $\Gyr$, divided by the true action
(computed in the correct coordinate system).

@@

\begin{figure*}
\plotone{fig/schmactions_many_orbits.pdf}
\caption{We report the 95\uth{} minus 5\uth{} percentile of the error in the
measured action ($\Delta J_i$) from coordinate system errors for the thin,
thick, and halo orbits (Table~\ref{tab:orbits}). The {\em left}, {\em center},
and {\em right} panels show the result for $J_r$, $L_z$, and $J_z$,
respectively. The {\em upper} panels consider an offset in $z$ and the {\em
lower} panels consider an offset in $x$ (equivalently, an offset in the Solar
radius).}
\label{fig:many_orbit_wrong_ref}
\end{figure*}

\begin{figure}
\plotone{fig/schmactions_many_orbits_Jz_fun.pdf}
\caption{The fractional error in $J_z$ as a function of $J_z$ for a few
different offsets in $z$. We show this for a $z$ offset of $10, 50, 100\,\pc$
(red, cyan, green respectively). As before, the error ($\Delta J_z$) is the
95\uth{} minus 5\uth{} percentile of the distribution of observed $J_z$ values
over the course of the orbit. There are large errors for thin-disk like orbits
($J_z \sim 0.5\,\actunit$), even for a small midplane offset of $10\,\pc$. }
\label{fig:dJz_fun_Jz}
\end{figure}

>>>>

\begin{deluxetable*}{cccccccc}
\tablecaption{Description and names of the three orbits considered in this
work.\label{tab:orbits}}
\tablehead{\colhead{name} & \colhead{initial position} & \colhead{initial
velocity} & \colhead{$J_r$} &
\colhead{$L_z$} & \colhead{$J_z$} & \colhead{$z_{\text{max}}$} &
\colhead{$A_R$}\\ \colhead{ } &
\colhead{$\mathrm{kpc}$} & \colhead{$\mathrm{km}/\mathrm{s}$} &
\colhead{$\mathrm{kpc\,km\,s^{-1}}$} &
\colhead{$\mathrm{kpc\,km\,s^{-1}}$} & \colhead{$\mathrm{kpc\,km\,s^{-1}}$} &
\colhead{$\mathrm{kpc}$} & \colhead{$\mathrm{kpc}$}}
\startdata 
thin-disk & $(8, 0, 0)$ & $(0, -190, 10)$ & 40.3 & -1520 & 0.69 & 0.12 & 1.29 \\
thick-disk & $(8, 0, 0)$ & $(0, -190, 50)$ & 32.5 & -1520 & 23.0 & 0.85 & 1.19 \\ 
halo & $(8, 0, 0)$ & $(0, -190, 190)$ & 32.8 & -1520 & 529.1 & 6.16 & 2.34
\enddata
\end{deluxetable*}

@@

\begin{figure}
\plotone{fig/schmactions_Jz_hist.pdf}
\caption{A histogram of the observed values in $J_z$ for the thick-disk orbit
assuming a $z$ offset of $100\,\pc$. One can see that if the observed $z$
values have a bias (from e.g. an incorrectly computed midplane), then the
induced error distribution in $J_z$ is decidedly non-Gaussian. Therefore, any
sort of error propagation must take this into account. We also plot one half
the 95\uth{} percentile minus the 5\uth{} percentile of each distribution as a
horizontal arrow anchored on the true $J_z$ value. We call this $\Delta J_z$
and will use it (along with the similarly defined $\Delta J_r$ and $\Delta
L_z$) to empirically describe the error distribution. We see that $\Delta J_z$
roughly corresponds to the distance from the true $J_z$ value to one of the
modes of the error distribution.}
\label{fig:Jz_hist}
\end{figure}

@@

Empirically describing the error distribution shown in
Figure~\ref{fig:Jz_hist} is difficult. Of course, Gaussian summary statistics
are not adequate. We therefore elect to measure this error by computing one half the
95\uth{} percentile minus the 5\uth{} percentile of the distribution of
observed action values. We refer to this quantity as $\Delta J_i$ for each
action value. We plot this quantity in Figure~\ref{fig:Jz_hist} as a horizontal arrow anchored on the true action value. Because of the bimodality of the error distribution, this
quantity roughly measures the distance from the true action value to the peak of one of the modes.
Furthermore, this bimodality also implies that $\Delta J_i$ is not very
sensitive to the exact percentiles used. This summary statistic does not
reflect the bias induced when the midplane error is $\gtrsim z_{\text{max}}$.

We now repeat the same procedure as in Figure~\ref{fig:one_orbit_wrong_ref}
but for systematic offsets between $0$ and $500\,\pc$ in the $z$ and $x$
components. In Figure~\ref{fig:many_orbit_wrong_ref}, we report one half the 95\uth{}
minus 5\uth{} percentile divided by the true action value ($\Delta J_i/J_i$) for the three
different fiducial orbits in Table \ref{tab:orbits}. The thick-disk
(\thickcolor) orbit is the one from Figure~\ref{fig:one_orbit_wrong_ref}, but
we also now consider the effect on the action determined for the thin-disk
(\thincolor) and halo (\halocolor) orbits.

The top row of Figure~\ref{fig:many_orbit_wrong_ref} shows the spread induced
in each action for an offset in the $z$-component (i.e. the Galactic
midplane). In the {\em bottom row} we consider offsets in the $x$ component
(i.e. the Solar radius). The {\em left}, {\em middle}, and {\em right} columns
show the fractional spread in the values computed for $J_z$, $L_z$, and $J_r$,
respectively. We compute the fractional spread $\Delta J_i/J_i$ as one half the 95\uth{}
minus 5\uth{} percentile of the action computed using the offset coordinate
system over the course of the first $\Gyr$, divided by the true action
(computed in the correct coordinate system).

@@

\begin{figure*}
\plotone{fig/schmactions_many_orbits.pdf}
\caption{We report one half the 95\uth{} minus 5\uth{} percentile of the error in the
measured action ($\Delta J_i$) from coordinate system errors for the thin,
thick, and halo orbits (Table~\ref{tab:orbits}). See discussion in the text for the justification in using this to measure the magnitude of the induced error. The {\em left}, {\em center},
and {\em right} panels show the result for $J_r$, $L_z$, and $J_z$,
respectively. The {\em upper} panels consider an offset in $z$ and the {\em
lower} panels consider an offset in $x$ (equivalently, an offset in the Solar
radius).}
\label{fig:many_orbit_wrong_ref}
\end{figure*}

\begin{figure}
\plotone{fig/schmactions_many_orbits_Jz_fun.pdf}
\caption{The fractional error in $J_z$ as a function of $J_z$ for a few
different offsets in $z$. We show this for a $z$ offset of $10, 50, 100\,\pc$
(red, cyan, green respectively). As before, the error ($\Delta J_z$) is one half the
95\uth{} minus 5\uth{} percentile of the distribution of observed $J_z$ values
over the course of the orbit. There are large errors for thin-disk like orbits
($J_z \sim 0.5\,\actunit$), even for a small midplane offset of $10\,\pc$. }
\label{fig:dJz_fun_Jz}
\end{figure}
